---
poster_height: "16.53in"
poster_width: "11.69in"
main_textsize: "42px"
body_textsize: "14px"
caption_textsize: "14px"
reference_textsize: "14px"
author_textsize: "20px"
authorextra_textsize: "14px"
affiliation_textsize: "14px" 
main_topsize: 0.1 #percent coverage of the poster
main_bottomsize: 0.04
main_fontfamily: Palatino
#ESSENTIALS
title: ''
author:
  - name: '**First Author (670000131)**'
    affil: 1
    main: true
  - name: Dr. Abu Bhakar
    affil: 2
  - name: Prof. Julie Thacker
    affil: 1
affiliation:
  - num: 1
    address: University of Exeter Medical School, Exeter, United Kingdom
  - num: 2
    address: Lalibela Hospital, Amhara, Ethiopia 

main_findings:
  - "   Opportunistic infections in HIV: <br>     diagnostics, management & outcomes in rural Ethiopia"
logoleft_name:  https&#58;//www.exeter.ac.uk/departments/communication/communications/design/downloads/logos/white_logo.png
column_numbers: 2
output: 
  posterdown::posterdown_betterport:
    self_contained: false
    pandoc_args: --mathjax
    number_sections: false
bibliography: references.bib
csl: "vancouver.csl"
link-citations: true
---

```{r, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      warning = FALSE,
                      tidy = FALSE,
                      message = FALSE,
                      fig.align = 'center',
                      out.width = "100%")
options(knitr.table.format = "html") 

library("dplyr")
library("ggplot2")
library("ggraph")
library("tidygraph")
```

# Introduction

Human immunodeficiency virus (HIV) prevalence in Ethiopia is 1.0 
(95% credibility interval 0.7-1.4) per 100,000 population; ten-fold higher 
than the UK [@unaids_unaids_2019].  
In Amhara region, north-eastern Ethiopia, 
1.2% (95%CI 0.7-1.4) of people had HIV in 2016 [@kibret_trends_2019].  
While antiretroviral therapy (ART) has improved outcomes in most countries; 
economically deprived rural areas still struggle with treatment retention and 
viral suppression. Thus immunocompromised patients suffer from infections they 
otherwise would not; contributing to a higher rate of mortality.  
Lalibela hospital in Amhara has some diagnostic laboratory capacity and an 
outpatient HIV department delivering ART. It is unclear whether there are 
facilities for opportunistic infections in advanced HIV 
infection (CD4 cell count <200/ml) [@assefa_outcomes_2011].  
This project aims to understand local hospital capacity in addressing diagnosis 
and management of opportunistic infections among HIV infected people,
as well as patient outcomes.  

## Objectives

1. Describe HIV patient care pathway in the Lalibela healthcare system. 
2. Describe diagnostic and management facilities in Lalibela hospital. 
3. Describe outcomes among patients in care. 

# Methods

This project involved a mixed methods approach, incorporating qualitative and 
quantitave aspects.  
The qualitative approach involved guideline review and semi-structured interviews
with local health care providers.  
The qualitative approach involved gathering health system indicators in the form 
of aggregate counts. This indicator data was gathered during the semi-structured 
interviews and included:  

- _Total catchment population_  
- _Number HIV cases in hospital cohort_  
- _Number cases with advanced HIV (CD4 cell count < 300 cells/mm^3^)_ 
- _Number of patients with laboratory tests for opportunistic infections_ 
- _Number of patients testing positive for opportunistic infections_ 
- _Number of patients virally suprssed_ 


# Results

**Patient care pathway**  
  
A decentralised healthcare system serving an ~350,000 catchment population; with 5 community health 
centres reporting to Lalibela hospital and 25 health posts reporting to each health centre, Figure \@ref(fig:healthsystem).  

- All hospital departments offer HIV RDT (high suspects tested) 
- Health posts also have HIV RDT testing 
- Initial and annual follow-up at hospital
  + CD4, Viral load, malnutrition screening 
- Medication dispensed at health centres 
  + Ease of access leading to better conitnuity of care
- All health care free under national insurance 
- National HIV programme supported by the Global Fund for HIV, TB and Malaria 

```{r, healthsystem, fig.cap='Visualisation of the health system in the Lalibela hospital catchment area', out.width="80%"}

points <- tibble(from = sort(c(rep.int(1, 5), 
                rep.int(2:6, 25))), 
                to = c(2:6, rep.int(NA, 125)), 
                health_type = c(rep.int("Hospital", 5), 
                         rep.int("Health centre", 125)))

hc <- 2:6

for (i in hc) {
  max_val <- max(points$to, na.rm = TRUE)
  
  points[which(points$from == i), 
         "to"] <- (max_val + 1):(max_val + 25)
}


graph <- as_tbl_graph(points)

ggraph(graph, layout = "partition", circular = TRUE) + 
  geom_node_arc_bar(aes(fill = factor(depth)), size = 0.25) + 
  coord_fixed() +
  scale_fill_manual(values = c("#0b4545", 
                               "#008080",
                               "grey95"), 
                    labels = c("Lalibela hospital", 
                               "Health centres (n = 5)", 
                               "Health posts (n = 125)")) + 
  theme_void(base_size = 18) + 
  theme(legend.title = element_blank()) 

```
<br> 
<br> 

**Diagnostics & management**  
  
There is a clear patient management pathway defined by a national programme.  
The hospital has well resourced laboratory capacity, which is supported by 
the US Centre for Disease Control and Prevention as well as Tulane 
University. The available lab test and management at the hospital level are displayed 
in Table \@ref(tab:labtable).

- Viral load and HIV resistance testing sent to capital lab (25 days for results)  
- CD4 count tests available at hospital 
- All patients presenting with CD4 count <350 prophylactically given broad spectrum ABx against diarrhoea and fungal infections

```{r, labtable}

tibble(Disease = c("Tuberculosis", 
                   "Sexually Transmitted Infections", 
                   "Crytococcal Meningitis", 
                   "Viral Hepatitis (B & C)"),
       Diagnostics = c("Microscopy <br> PCR", 
                       "ELISA", 
                       "Culture", 
                       "PCR <br>
                       ELISA"), 
       Management = c("Negative: prophylatic Isoniazide <br> Positive: Tripple therapy", 
                      "Broad spectrum ABx",
                      "Appropriate ABx", 
                      "Patient pays privately for test, <br> 
                      managment privately in capital"
                      )) %>% 
  knitr::kable(
  format = "html",
  caption = "Diagnostic and management facilities for HIV associated OIs",
  align = "c",
  escape = FALSE)
```


**Outcomes**  
  
With the national pharmacy association supplying hospitals, there are infrequent 
ruptures. However there are shortages of second line ARV. Despite this there are
quite impresive outcomes as seen in Figure \@ref(fig:cascade).  

- 88% of those in the cohort are virally surpressed 
- Among those with OIs: 55 had diarrhoea or skin infections, only 1 patient with extrapulmonary TB  
  
  
```{r cascade, fig.cap='Cascade of care for opportunistic infections among HIV infected patients <br> (N = 350,000)', fig.height=6, out.width="80%"}

# create a dataframe with counts for plotting 
numbers <- tidyr::tibble(measure = c("HIV cases", 
                                     "Advanced \n HIV cases \n (CD4 < 300)", 
                                     "Advanced \n HIV cases tested \n for OIs",
                                     "Cases with OIs"), 
                         countz = c(1502, 
                                    231, 
                                    231, 
                                    55)
                         ) %>% 
  # create proportions, dividing by previous row
  mutate(measure = factor(measure, measure), 
         prop = round(countz / lag(countz) * 100, digits = 1), 
         prop = tidyr::replace_na(prop, 100))


numbers[which(numbers$measure == "HIV cases"), "prop"] <- 0.4

numbers <- mutate(numbers, 
                  # pull together labels of counts and props
                  labz = glue::glue("{countz} \n ({prop}%)")
                  )
         


# initiate plot 
ggplot(data = numbers) + 
  # plot stacked bar charts with number of visits categorised by tested or not 
  geom_bar(aes(x = measure, y = countz), fill = "#0b4545", stat = "identity") + 
  # flip to make horizontal
  coord_flip() + 
  # add proportions as text labels 
  geom_text(aes(x = measure, y = countz , label = labz), hjust = -0.15) +
  # fix y axis plotting 
  scale_x_discrete(limits = rev(levels(numbers$measure))) + 
  # fix x axis plotting 
  scale_y_continuous(expand = c(0,0), 
                     limits = c(0, 2000), 
                     breaks = seq(0, 2000, 250)) + 
  # make classic bar chart and increase text size 
  theme_classic(base_size = 18) + 
  # make x axis labels slanted, remove legend title and move to bottom 
  theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1), 
        legend.title = element_blank(), 
        legend.position = "bottom", 
        legend.box = "vertical") + 
  # change the axis titles 
  labs(x = "", y = "Counts (n% of above group)")



```

# Discussion / Conclusion

The estimated HIV prevalence based on the hospital catchment population is much lower
than literature described values for the region.  
This along with the annecdotal reasoning that Lalibela prefecture is better off 
than the surrounding region may explain the very impressive outcomes.  
Nevertheless this appears to be a well structured and resourced health system 
for addressing opportunistic infections in HIV patients. 
Future studies should consider the impact on TB monotherapy on antimicrobial resistance
prevalence, as well as the potential impact of making viral hepatitis testing and 
management locally available free of charge. 

# References
